### 2025-08-04 mixed-models.qmd
20250804 - comparing model fit

## Data Structure {#sec-mixedModelDataStructure}

@@

@@

We compare models using the `bbmle` [@R-bbmle] and `MuMIn` [@R-MuMIn] packages.

```{r}
anova(
  pointsPerSeason_linearRegression,
  pointsPerSeason_quadraticRegression
)

@@

```{r}
anova(
  pointsPerSeason_randomIntercepts,
  pointsPerSeason_position
)

anova(
  pointsPerSeason_positionAgeFixedLinearSlopes,
  pointsPerSeason_positionAgeRandomLinearSlopes,
  pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes,
  pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction
)
```

@@

##### Generalized Additive Model {#sec-fantasyPointsByAgeExperienceModelsAgeGAM}

@@

###### Compare Models {#sec-fantasyPointsByAgeExperienceModelsAgeGAMcompare}

>>>>

### Data Structure {#sec-mixedModelDataStructure}

@@

### Comparing Model Fit {#sec-mixedModelCompareModels}

When trying to determine what predictors to include and how complex of a model to fit, it can be helpful to compare a variety a models to determine how well they fit the data (i.e., how much variance they account for).
To compare non-nested models, you can use fit criteria such as Akaike information criterion (AIC), which penalizes model complexity.
If two models are nested, you can compare their model fit using a likelihood ratio test.
Two models are considered nested if one model includes all of the terms of the original model along with additional terms.
The simpler model is considered "nested within" the more complex model.
In the likelihood ratio test, a significant difference indicates that the more complex model is significantly better fitting than the simpler model.
If the likelihood ratio test is non-significant, it indicates that the more complex model is not significantly better fitting, and thus you should prefer the more parsimonious model.

@@

We compare nested models using an $F$-test with the `anova()` function.
We compare non-nested models using the `bbmle` [@R-bbmle] and `MuMIn` [@R-MuMIn] packages.

```{r}
anova(
  pointsPerSeason_nullModel,
  pointsPerSeason_linearRegression,
  pointsPerSeason_quadraticRegression
)

@@

```{r}
anova(
  pointsPerSeason_randomIntercepts,
  pointsPerSeason_position,
  pointsPerSeason_positionAgeFixedLinearSlopes,
  pointsPerSeason_positionAgeRandomLinearSlopes,
  pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes,
  pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction
)
```

@@

#### Generalized Additive Model {#sec-fantasyPointsByAgeExperienceModelsAgeGAM}

@@

##### Compare Models {#sec-fantasyPointsByAgeExperienceModelsAgeGAMcompare}
